Comparison of energy costs of commuting in England and the Netherlands
========================================================

Before attempting to run these commands, you should have first downloaded and unzipped the 'energy-nat' folder onto your computer (download it [here](https://dl.dropboxusercontent.com/u/15008199/egs2stay/energy-nat.zip)) and have processed the [English](http://rpubs.com/RobinLovelace/7178) and [Dutch](http://rpubs.com/RobinLovelace/7180) commuting data examples. The resulting objects need to be on your R workspace when you run through this code, as it refers to data loaded from these sessions. 

Preliminary analysis of English energy costs
-----
```{r}
load("en-nl.RData")
gors$label <- gortex$V3 # set the zone codes to identify the English zones
gors$ZONE_CODE <- gortex$V3 # again, for reference
gors$EAV <- gors$ET/gors$all.all # average energy costs: total divided by the population 
gors@data[,c(1:5,115,114, 116)]
summary(gors$EAV) # browsing out dataset
sum(gors$ET) / sum(gors$all.all) # EN population-weighted average, lower than average of GORs
```

Preliminary analysis of NL energy costs
-----
```{r}
r3$pops <- c(2415, 1180, 3530, 2583, 482, 642, 1968, 576, 1143, 1105, 356, 378  )
sum(r3$pops) # Adding populaitons of NL provinces to calculate per person averages
summary(r3$etot)
sum(r3$Totaal.vervoerwijzen * r3$etot)/((mean(r3$Totaal.vervoerwijzen))*12) # Netherlands region average (unweighted)
r3@data[,c("name", "code_hasc", "Vervoerwijzen", "Totaal.vervoerwijzen", "etot", "pops")] # re-check data
r3$etot * r3$pops
sum(gors$ET) / sum(gors$all.all) # En pop weighted
sum(r3$etot * r3$pops) / sum(r3$pops) # NL pop-weighted av. also lower, due to cities
38.1/34.5 # 10% difference...
(sum(gors$ET) / sum(gors$all.all)) / (sum(r3$etot * r3$pops) / sum(r3$pops))
```
Comparing the modal split of commuter trips in England and the Netherlands
---------------------
```{r}
msplit.nl <- r3@data[,7:14]
for(i in 1:8){
  msplit.nl[,i] <- r3@data[,i+6] * r3$pops
}
m.nl <- colSums(msplit.nl)
names(m.nl)[4] <- "Metro"

m.nl <- m.nl * 500
m.nl <- data.frame(tot = m.nl, prop=m.nl/sum(m.nl) * 100) ; m.nl$mode <- row.names(m.nl)
m.nl <- m.nl[-c(3,5,8),]
m.nl$Country <- "Netherlands"

head(gors@data)
m.en <- colSums(gors@data[,c(5,6,8,9,11,12,4)]) / sum(gors$all.all)
m.en <- colSums(gors@data[,c(4:13)])
m.en <- data.frame(tot = m.en, prop = m.en/sum(m.en) *100)
wones <- c(4,6,8,9,11,12) - 3
m.en <- m.en[wones,]
m.en$mode <- c("Metro", "Bus", "Car.d", "Car.p", "Bicycle", "Walk")
m.en$Country <- "England"

m.all <- rbind(m.en, m.nl)
head(m.all)
p <- ggplot(m.all, aes(y = prop, fill = mode))
p + geom_bar(aes(x=mode), stat="identity") + facet_grid(facets= . ~ Country) + 
  scale_fill_brewer(type="qual", palette=7) +
  ylab("% of commuter trips")
```

Comparing distances
-----------
```{r}
names(gors)[4:15]
md # mean distances, saved from "english analysis"
md <- md[,-1]
t.en <- colSums(gors@data[,3:101])
md
md$mdtotald <- 0
for(i in 1:11){
  md[i,"mdtotald"] <- sum(md[i,1:9] * colSums(gors@data[,seq(from=4, by=11, length.out=9) + i-2]))
}
md

md$dist <- md$mdtotald / t.en[1:11]
md$pdist <- md$mdtotald / sum(md$mdtotald)
md
barplot(md$mdtotald, col=1:11)
md$tdist <- md$mdtotald ; md <- md[,-10]
md2 <- md[c(3,4,6,7,9,10,2),]
md2
md2$Country <- "England"
sum(md$tdist) / sum(gors$all.all) # Total distance by all modes: 10 km

# Distance NL
names(r3)
md.nl <- data.frame(dist = colSums(r3@data[,16:23]) / nrow(r3@data))
md.nl # v. useful infor: compare w.
md.nl$tdist <- md.nl[,1] * colSums(r3@data[,7:14])
md.nl$pdist <- md.nl$tdist / sum(md.nl$tdist)
md.nl$Country <- "Netherlands"
qplot(data=md.nl, x=row.names(md.nl), y=pdist)
md.nl$ptrips <- colSums(r3@data[,7:14]) / nrow(r3@data) 

sum(md.nl$dist * md.nl$ptrips) 
md.nl

md.all <- rbind(md2[,c("dist", "pdist", "tdist", "Country")], 
                md.nl[-which(row.names(md.nl) == "Moto.1" | row.names(md.nl) == "Other.1"),
                      c("dist", "pdist", "tdist", "Country")])
md.all
md.all$mode = c("Train", "Bus", "Car.d", "Car.p", "Bicycle", "Walk", "Metro", 
                "Car.d", "Car.p", "Train", "Metro", "Bicycle", "Walk")
md.all

qplot(data=md.all[-c(1,10),], x=mode, y=dist, fill=mode, geom="bar") + facet_grid(facets= . ~ Country) + 
  scale_fill_brewer(type="qual", palette=7) +
  ylab("Distance of commute (km)")

md.all$ptrips <- 0
md.all$ptrips[8:13] <- colMeans(r3@data[,c(1,2,3,4,6,7)+6])
md.all$ptrips[1:7] <- colSums(gors@data[,c(5,6,8,9,11,12,4)]) / sum(gors$all.all)
md.all
```

Basic analysis of average distances
--------
```{r}
sum(md.all$dist[1:7] * md.all$ptrips[1:7]) # 14.5 av dist uk
sum(md.all$dist[8:13] * md.all$ptrips[8:13])
md.all$dist[md.all$mode == "Car.d"][2] / md.all$dist[md.all$mode == "Car.d"][1]
```

The relationship between energy use and population density
-------
```{r}
library(rgeos)
gors$area <- gArea(gors, byid=T) / 1000000
r3$area <- gArea(r3, byid=T) / 1000000
r3$popdens <- r3$pops * 1000 / r3$area
gors$popdens <- gors$all.all * 2 / gors$area
r3$EAV <- r3$etot

pden.all <- rbind(r3@data[,c("popdens", "EAV")], gors@data[,c("popdens", "EAV")])
pden.all$Country <- c(rep("Netherlands", 12), rep("England", 9))
qplot(data=pden.all, x=popdens, y=EAV, color=Country) + ylab("Energy costs of commuting (MJ/trip)") +
  xlab("Population density (ppl/sqkm)") + geom_abline(intercept=sum(gors$ET) / sum(gors$all.all),slope=0, colour="red") + 
  geom_abline(intercept=sum(r3$etot * r3$pops) / sum(r3$pops),slope=0, colour="blue") 
```



