Loading and plotting English commuting data 
========================================================

This is document is based on "england.R", a script file that loads raw data on commuting in England, analyses it, and then plots it at the regional level. This is the first of three R script files for comparing the energy costs of commuting in the England and the Netherlands. (The other two can be found [here](http://rpubs.com/RobinLovelace/7182) and [here](http://rpubs.com/RobinLovelace/7180)). The input data and code is a available as a [.zip file](https://dl.dropboxusercontent.com/u/15008199/egs2stay/energy-nat.zip). The R packages we will be using to do the analysis include ggplot2 (for plotting), sp (for spatial data analysis) and rgeos for geographic functions. Load them in batch with the following two commands:

```{r}
x = c("ggplot2", "sp", "rgeos", "mapproj", "rgdal", "maptools")
lapply(x, require, character.only=T) # the R packages we'll be using
```

As with most data analysis tasks, the first stage is to load and view the data.

Loading and viewing the English commuting data
-----------
Set R's working directory to the right place for the English data (the folder "EN-Data"). For this to work, R must already have "energy-nat" as its working directory. Rstudio users can do this by clicking 'Session > Set Working Directory > To Source file location'. The shapefile data is loaded using the command readOGR:

```{r}
setwd("EN-Data/") # navigate to the correct working directory
gors <- readOGR(".", layer="GOR_st121" ) # load the data!
```

Now that the English regions are loaded, it's time to naviagate back to the previous directory, take a look at the data we have loaded, abbreviate the region names, and plot the results to ensure they make sense:

```{r}
setwd("../")
getwd() # check we're in the correct directory: energy-nat
class(gors) # check we've got the right R object: SpatialPolygonsDataFrame
gors@data[,1:5] # take a look at the first 5 columns of data: integer counts
gors$abbed <- abbreviate(gors$ZONE_LABEL, minlength=2) # create abbreviated area names
gors@data[1:5,c("ZONE_LABEL", "abbed")] # check the abbreviation worked
plot(gors) # results in a plot of England
text(coordinates(gors), labels=gors$abbed, col="red") # add our region names: in the right place?
```

Adding distances and calculating energy costs for modes
---------------
Now we need to make sense of the 103 columns of data that constitute the regional data ("gors@data"). At the moment the columns have names such as ST121001. We need to rename them so they make sense as commuting modes and distance bands:
```{r}
dbreaks <- c(2,5,10,20,30,40,60) # these are the distance breaks in the data
dbreaks2 <- c("all", 0, dbreaks) # correct labels to use with our data
# other than 002 it's an 11 by 9 matrix
modenames <- c("all", "metro", "train", "bus", "moto", "car", "carp", "taxi", "bike", "walk", "other")
newnames <- paste(rep(modenames, times=9), rep(dbreaks2, each=11), sep=".") # give each column a unique name
newnames[1:20] # Shows the logical structure of the column names created with the above command
gors@data$mfh <- gors@data$ST1210002 # rename the column that does not fit: mainly from home (mfh)
gors@data <- gors@data[,-4] # remove the original mainly working from home column: gets in the way
names(gors@data)[3:101] <- newnames
ecosts <- c(0,0.57,0.77,2.1,1.7,3,0,3,0.09,0.13,0) # these are the direct energy costs (MJ/KM) for each mode
names(ecosts) <- c("All", "Tram", "Train", "Bus", "Moto", "Car.d", "Car.p", "Taxi", "Cycle", "Walk", "Other")
barplot(ecosts[2:9]) # Make sense?
```
The data we now have is counts of people travelling by certain modes in specific distance boundaries. The problem is that these distance bins apply for Euclidean (straight line) distance. To calculate energy costs, we need to know the average route distance. To solve this problem, an external dataset (Understanding Society) was used to identify the most common route distances within each category; the average [circuity](http://en.wikibooks.org/wiki/Transportation_Geography_and_Network_Science/Circuity) factor is assumed to be 1.4 ([Ballou, 2002](http://www.sciencedirect.com/science/article/pii/S0965856401000441)). This stage introduces much uncertainty: we cannot be sure that the route distance for each distance boundary are as illustrated below, and more research is clearly needed in this area.
```{r}
md <- read.csv("dbxs.csv", header=T, ) # the distances associated with each distance bin
plot(rep(c(0,0,1,3.5,7.5,15,25,35,50,100),11), t(md)) # visualise them: make sense?
```
With estimates of route distances for different modes now saved, we can progress to calculate the energy costs of each distance/mode category.

```{r}
mdv <- as.vector(as.matrix(md[,2:10])) # rearange the distances data
edcosts <- rep(ecosts, 9) * mdv # multiply the energy costs by average distances fo each of 9 modes
ecs <- sweep(x=gors@data[,3:101], MARGIN=2, edcosts, '*' ) # multiply the count data in each cell by their respective energy cost, for every zone
(ecs)[seq(6,by=11,length.out=9)] # test for one mode: cars (for all zones)
ecs.mode <- cbind(rowSums(ecs[,seq(2,by=11,length.out=9)]))
ecs.mode <- data.frame(array(dim = c(9,11)))
for(i in 1:11){
  ecs.mode[,i] <- cbind(rowSums(ecs[,seq(from=i,by=11,length.out=9)]))
}
paste(names(ecs),"en", sep=".")
names(ecs.mode) <- paste("moden",modenames, sep=".")
gors@data <- cbind(gors@data, ecs.mode)
gors@data$ET <- rowSums(ecs)
plot(gors, col = rowSums(ecs))
```

Save the results and prepare for plotting
-------------------

```{r}
# writeOGR(obj=gors, dsn=".", layer="gorens", driver="ESRI Shapefile") # uncomment to save
# convert to ggplot2, just for plotting
fgor <- fortify(gors, region="ZONE_LABEL")
fgor <- merge(fgor, gors@data[,c(1,2,3,8,ncol(gors@data))], 
            by.x = "id", by.y="ZONE_LABEL")
head(fgor)
gortex <- coordinates(gors)/1000 
gortex <- as.data.frame(cbind(gortex))
gortex$V3 <- gors$abbed
head(gortex)
class(gortex$V1)
```

Plotting the final results
----------
Finally, we use ggplot2 as a static GIS, to plot the results in a certain way (to be comparable with the Nethelands data):
```{r}
p <- ggplot(data=fgor, aes( x=long/1000, y=lat/1000))
p + geom_polygon(data=fgor,aes(x=long/1000, y=lat/1000, fill=ET/all.all, group=group)) + 
  geom_path(data=fgor,aes(x=long/1000, y=lat/1000,group=group)) +
  scale_fill_continuous(low="green", high="red", name="Etrp (MJ)",
                        limits=c(20,55)) + coord_fixed() +
  xlab("Easting (km)") + ylab("Northing (km)") +
  geom_text(data=gortex, aes(x=V1, y=V2, label=V3), colour="blue") #+ theme_minimal()
```


